Learning how to act: making good decisions with machine learning

### 2017-07-09 thesis/thesis_map.tex
rough draft of chapter 4

Stratified randomised experiments address the issue of variance due to covariate imbalance by randomly allocating treatment conditional on covariates belived to influence the outcome of interest. This introduces a backdoor path from treatment to outcome via the covariates on which treatment is stratified, see figure \ref{fig:random_experiment_network_stratified}, nessessitating that one condition on these covariates in computing the average casual effect in the same way as for estimating causal effects under under ignorability \S\ref{sec:causal_effect_under_ignorability}. The key difference is that the propensity score is known, as it is designed by the experimenter, and there are guaranteed (rather than assumed) to be no latent confounding variables (that influence both treatment and outcome). See \citet{imbens2015causal} for a discussion of the trade-offs between statified versus completely random experiments.

IF THE PROPENSITY SCORE IS THE SAME FOR ALL COVARIATES IT DOESN'T SEEM ONE MUST ADJUST.
\todo{double check}

\begin{figure}
\caption{causal network for a stratified randomised experiment}
\label{fig:random_experiment_network_stratified}
\centering
\begin{tikzpicture}[->,shorten >=0pt,shorten <=0pt,node distance=3em,thick, node/.style={observedrect}, lt/.style={latent}]
\node[node](1){Treatment};
\node[node, above left=of 1](2){Randomiser};
\node[node, above right=of 1](5){Covariates};
\node[node, below right=of 5](3){Outcome};
\node[lt, above=of 3](4){U};

@@

\todo{discuss the key ways in which the simple regret problem differs from cummulative regret}

@@

It is worth noting that definition \ref{def:contextual-bandit-problem} does not make any assumptions about the \emph{causal} relationship between the context $\vec{X}$ and the reward $\vec{Y}$, see figure \ref{fig:contextual-bandit-causal-graph}. \citet{Bareinboim2015} demonstrate that policies that incorporate observations of the action an agent would have taken were their action not set by the bandit policy can acheive lower regret than those that ignore this information. This is an example of the case represented in figure \ref{fig:context-graph-confounded}. However, the context should be relevent, such that $\P{\vec{y}|\vec{x}} \neq \P{\vec{y}}$, otherwise including it is the equivelent to adding irrelevent featurs to a supervised learning problem.

@@

\subsubsection{Learning from logged bandit data}
Another topic of interest within the contextual bandit community, which is deeply connected to causal effect estimation from observational data, is learning from logged bandit feedback data or off-policy evaluation\citep{Langford2008a,strehl2010learning,Li2010,Dudik2011,Bottou2013,Swaminathan2015}. In this setting, the learner has a data set $S = \set{(\vec{X}_t,A_t,\cfb{Y}{A_t})}_{t=1}^T$, which is assumed to have been generated by a stochastic contextual bandit enviroment interacting with some unknown, potentially stochastic, policy $\pi(\vec{x}_t,h_t)$, where $h_t$ is the sequence of observed data up to time $t$. The goal of the learner is to evaluate the value of an alternate policy, $\pi'$, for selecting actions, often with the underlying motivation of identifying an optimal policy within some space of policies $\Pi$.

@@

@@

A well-studied class of bandit problems with side information are ``contextual bandits''~\cite{Langford2008,Agarwal2014}. Our framework bears a superficial similarity to contextual bandit problems since the extra observations on non-intervened variables might be viewed as context for selecting an intervention. 
However, a crucial difference is that in our model the extra observations are only revealed \emph{after} selecting an intervention and hence cannot be used as context.

>>>>

Stratified randomised experiments address the issue of variance due to covariate imbalance by randomly allocating treatment conditional on covariates belived to influence the outcome of interest. If we stratify in such a way that the probability an instance recieves a given treatement is indepdenent of its covariates, the causal graphical model in figure \ref{fig:random_experiment_network} still holds and we can still estiamte the average causal effects directly from the differences in outcome accross treatments. However, more complex stratification strategies can introduce a backdoor path from treatment to outcome via the covariates on which treatment is stratified, see figure \ref{fig:random_experiment_network_stratified}, nessessitating that one condition on these covariates in computing the average casual effect in the same way as for estimating causal effects under under ignorability \S\ref{sec:causal_effect_under_ignorability}. The key difference is that the propensity score is known, as it is designed by the experimenter, and there are guaranteed (rather than assumed) to be no latent confounding variables (that influence both treatment and outcome). See \citet{imbens2015causal} for a discussion of the trade-offs between statified versus completely random experiments.

\begin{figure}
\caption{causal network for a stratified randomised experiment if the probability an individual is assigned a given treatment depends on some covariates.}
\label{fig:random_experiment_network_stratified}
\centering
\begin{tikzpicture}[->,shorten >=0pt,shorten <=0pt,node distance=3em,thick, node/.style={observedrect}, lt/.style={latent}]
\node[node](1){Treatment};
\node[node, above left=of 1](2){Randomiser};
\node[node, above right=of 1](5){Covariates};
\node[node, below right=of 5](3){Outcome};
\node[lt, above=of 3](4){U};

@@

%\todo{discuss the key ways in which the simple regret problem differs from cummulative regret}

@@

It is worth noting that definition \ref{def:contextual-bandit-problem} does not make any assumptions about the \emph{causal} relationship between the context $\vec{X}$ and the reward $\vec{Y}$, see figure \ref{fig:contextual-bandit-causal-graph}. However, the context should be relevent, such that $\P{\vec{y}|\vec{x}} \neq \P{\vec{y}}$, otherwise including it is the equivelent to adding irrelevent featurs to a supervised learning problem. \citet{Bareinboim2015} demonstrate that, in some cases, policies that incorporate observations of the action an agent would have taken were their action not set by the bandit policy can acheive lower regret than those that ignore this information. This is an example of the case represented in figure \ref{fig:context-graph-confounded}.

@@

\subsection{Learning from logged bandit data}
Another topic of interest within the bandit community, which is deeply connected to causal effect estimation from observational data, is learning from logged bandit feedback data or off-policy evaluation\citep{Langford2008a,strehl2010learning,Li2010,Dudik2011,Bottou2013,Swaminathan2015}. In this setting, the learner has a data set $S = \set{(\vec{X}_t,A_t,\cfb{Y}{A_t})}_{t=1}^T$, which is assumed to have been generated by a stochastic contextual bandit enviroment interacting with some unknown, potentially stochastic, policy $\pi(\vec{x}_t,h_t)$, where $h_t$ is the sequence of observed data up to time $t$. The goal of the learner is to evaluate the value of an alternate policy, $\pi'$, for selecting actions, often with the underlying motivation of identifying an optimal policy within some space of policies $\Pi$.

@@

\subsection{Adding structure to actions}

The classic multi-armed bandit is a powerful tool for sequential decision making. However, the regret grows linearly with the number of (sub-optimal) actions and many real world problems have large or even infinite action spaces. This has led to the development of a wide range of models that assume some structure across the reward distributions for different arms, for example generalised linear bandits \citep{filippi2010parametric}, dependent bandits \citep{Pandey2007}, X-armed bandits \citep{Bubeck2010} and Gaussian process bandits \citep{srinivas2009gaussian}, or that consider more complex feedback, for example the recent work on graph feedback\citep{Mannor2011,Lelarge2012,Alon2013,Buccapatnam2014,Kocak2014,Alon2015} and partial monitoring \citep{Piccolboni2001,Bartok2014}.

In the next chapter, I propose a very natural connection between causal graphs and bandit problems and show it induces a novel form of additional feedback and structure between arms that cannot be can exploited by any of these previous approaches.

@@

Our framework bears a superficial similarity to contextual bandit problems, \S\ref{sec:contextual-bandits}, since the extra observations on non-intervened variables might be viewed as context for selecting an intervention. 
However, a crucial difference is that in our model the extra observations are only revealed \emph{after} selecting an intervention and hence cannot be used as context.
